\LARGE \bf Thymio, an example of open source hardware robot \\ for formal education

### 2016-03-29 paper_ram.tex
small modifs

There is another licensing issue which is not well known and that we discovered very late in our project: the constrains of the license of the mechanical and electronic CAD tools. 
Indeed, asked about this issue, people in the field seems generally not aware of the fact that CAD license can be very restrictive in the way the source files can be published.
Around the half or respondents to our survey, also among the project leaders, state that they were not aware that this could be an issue, and only one third checked the license of their CAD software. 
Among the respondents, some stated that they did not checked the license because they were sure they can freely publish their design, but half of them are using software not allowing the publication of source files when using educational licenses, and all these respondents are academic or students.
This issue is very serious, as most contributors of open source hardware projects are academic and use academic or educational licenses. 
Most licenses do not allow to use the design created with this version of the software for commercial activities. 
As producing or selling the product is part of the definition of open hardware, these licenses simply forbid publication under the standard open source hardware conditions.

@@

To clarify this issue we contacted twelve of the major editors of mechanical CAD and PCB routing software.
We asked them if their educational license allows the publication of the source files, and we specified that the published files could have been downloaded by a company to produce the system.
Figure~\ref{fig:editors} summarize the result of this survey.
Only three of the twelve editors have licenses allowing this type of publication. 
Two others explicitly mentioned the possibility to ask for permission before publication.
A large mechanical CAD editor was puzzled by our questions and after realizing the impact of the license, introduced a special condition allowing publication of files in clearly labeled open source hardware projects.
In previous situations of open source publication, the same editor asked the universities to purchase commercial licenses to permit publication.
This can multiply by a factor of some hundreds the price of the CAD license.
This blocking factor for open hardware also apply to the publication of scientific results, as promoted by many governments in the last decade and generally called ``open science''.
The change of policy in CAD files publication that we reached is a sign of the very positive trend set by open hardware and by open science in general.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/table}
\caption{Publication possibilities as function of the CAD editors.}
\label{fig:editors}
\end{figure}

As a conclusion, this issue is totally underestimated by both people participating to the projects and by the CAD editors. 
In a period where editors are looking for additional revenues and are attacking universities for abuse of licenses\footnote{We are aware, just in Switzerland, of two situations where large amounts of money are asked to universities for non-respect of software licenses.}, this can be a very dangerous situation.

@@

Among the fundamental choices when starting an open source hardware project, there is the type of production. 
In the definition of open hardware we gave, it is stated that we should offer an ``hardware whose design is made publicly available so that anyone can ... make... the ... hardware ''.
Behind the ``anyone'' should we consider every single person or only companies able to produce the product?
This choice has implications on who designs the system and how.
In our project we have two different types of hardware: the robot itself and the accessories used in specific activities.
The robot is the expensive part and has a very neutral design, allowing adaptation to specific situations.
This adaptation is achieved by custom accessories that increase the attraction of the robot in its specific role, enabling activity for different ages and genders.

>>>>

There is another licensing issue which is not well known and that we discovered very late in our project: the constrains of the license of the mechanical and electronic CAD tools. 
Indeed, asked about this issue, the participants to our survey seem generally not aware of the fact that CAD licenses can be very restrictive about the way the source files can be published.
Around the half or respondents to our survey, also among the project leaders, state that they were not aware that this could be an issue, and only one third checked the license of their CAD software. 
Among the respondents, some stated that they did not checked the license because they were sure they can freely publish their design, but half of them are using software not allowing the publication of source files when using educational licenses, and all these respondents are academic or students.
This issue is very serious, as most contributors of open source hardware projects are academic and use academic or educational licenses. 
Most licenses do not allow to use the design created with this version of the software for commercial activities. 
As producing or selling the product is part of the definition of open hardware, these licenses simply forbid publication under the standard open source hardware conditions.

@@

To clarify this issue we contacted twelve of the major editors of mechanical CAD and PCB routing software.
We asked them if their educational license allows the publication of the source files, and we specified that the published files could have been downloaded by a company to produce the system for commercial purposes.
Figure~\ref{fig:editors} summarizes the result of this survey.
Only three of the twelve editors have education licenses allowing this type of publication. 
Two others explicitly mentioned the possibility to ask for permission before publication.
A large mechanical CAD editor was puzzled by our questions and after realizing the impact of the license, introduced a special condition allowing publication of files in clearly labeled open source hardware projects.
In previous situations of open source publication, the same editor asked the universities to purchase commercial licenses to permit publication.
This can multiply by a factor of some hundreds the price of the CAD license.
This blocking factor for open hardware also apply to the publication of scientific results, as promoted by many governments in the last decade and generally called ``open science''.
The change of policy in CAD files publication that we obtained is a sign of the very positive trend set by open hardware and by open science in general.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/table}
\caption{Publication possibilities as function of the CAD editors, situation in February 2016.}
\label{fig:editors}
\end{figure}

As a conclusion, this legal issue is totally underestimated by both people participating to the projects and by the CAD editors. 
In a period where editors are looking for additional revenues and are attacking universities for misuse of licenses\footnote{We are aware, just in Switzerland, of two situations where large amounts of money are asked to universities for non-respect of software licenses.}, this can be a very dangerous situation.

@@

Among the fundamental choices when starting an open source hardware project, there is the choice of the type of production. 
In the definition of open hardware we gave, it is stated that one should offer an ``hardware whose design is made publicly available so that anyone can ... make... the ... hardware ''.
Behind the ``anyone'' should we consider every single person or only companies able to produce the product?
This choice has implications on who designs the system and how.
In our project we have two different types of hardware: the robot itself and the accessories used in specific activities.
The robot is the expensive part and has a very neutral design, allowing adaptation to specific situations.
This adaptation is achieved by custom accessories that increase the attraction of the robot in its specific role, enabling activity for different ages and genders.
